---
title: "exploratory data analysis"
output:
  html_document:
    theme: lumen
    #code_folding: hide
    df_print: paged
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
---

```{r}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggplot2)
library(gsubfn)
library(cowplot)
library(ggpubr)
options(warn=-1)
options(dplyr.summarise.inform = FALSE)

apatheme=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(color="black"),
        text=element_text(size=10),
        legend.title=element_blank())

apatheme_2=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(color="black"),
        axis.text.x = element_text(angle = 70, hjust = 1),
        text=element_text(size=10),
        legend.title=element_blank())
```

# load preprocessed data

```{r}
list_all_files <-list.files("./raw_data/main_experiment/",full.names = TRUE)
d<-do.call("rbind",lapply(list_all_files,read.csv, header=TRUE))

sum(is.na(d))
dim(d) # 33000    35
```

# removal of trials that did not meet the RT criteria

```{r}
before_removal <- nrow(d)
d<-d[d$rt > .1 & d$rt < 1.5,] # measured in seconds
nrow(d) # 31205 obs
number_removed <- (before_removal - nrow(d))
number_removed # 1795

# percentage of removed trials
(before_removal - nrow(d))/before_removal *100 # 5.43 %

dim(d)
```

# active trials count (per individual)

```{r}
# count the number of active trials per participant
same_agent_responses_count <- d %>%
  group_by(individual_id) %>%
  count()

# change to character
same_agent_responses_count$individual_id <- as.character(same_agent_responses_count$individual_id)

# bar plot for the active trials count
bar_plot <- ggplot(same_agent_responses_count, aes(x = individual_id, y = n)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of active trials per participant",
       x = "Participant ID",
       y = "Response Count") +
  apatheme_2

print(bar_plot)

mean_response_count <- mean(same_agent_responses_count$n)
sd_response_count <- sd(same_agent_responses_count$n)
range_response_count <- range(same_agent_responses_count$n)

# participants with < 500 active trails
participants_fewer_than_500 <- same_agent_responses_count %>%
  filter(n < 500)
print(participants_fewer_than_500)

# count trials of left/right responses per participant

# filter data for left responses
left_responses <- d %>% filter(response_n == "0")
left_responses_stats <- left_responses %>%
  group_by(individual_id) %>%
  summarise(count_leftward = sum(length(trial)))

# filter data for right responses
right_responses <- d %>% filter(response_n == "1")
right_responses_stats <- right_responses %>%
  group_by(individual_id) %>%
  summarise(count_rightward = sum(length(trial)))

mean_left <- mean(left_responses_stats$count_leftward)
range_left <- range(left_responses_stats$count_leftward)
sd_left <- sd(left_responses_stats$count_leftward)

mean_right <- mean(right_responses_stats$count_rightward)
range_right <- range(right_responses_stats$count_rightward)
sd_right <- sd(right_responses_stats$count_rightward)

print(mean_left) # 232
print(range_left) # 107 324
print(sd_left) # 39

# count the number of trials preceded by own/partner

# filter for trials preceded by own
preceeded_by_own_trial <- d %>%
  filter(individual_id == lag(individual_id))

# group by individual id
preceeded_by_own_trial_stats <- preceeded_by_own_trial %>%
  group_by(individual_id) %>%
  summarise(count_preceeded_by_own = n())

preceeded_by_own_trial_stats

# filter for trials preceded by partner
preceeded_by_partner_trial <- d %>%
  filter(individual_id != lag(individual_id))

# group by individual id
preceeded_by_partner_trial_stats <- preceeded_by_partner_trial %>%
  group_by(individual_id) %>%
  summarise(count_preceeded_by_partner = n())

preceeded_by_partner_trial_stats

mean_trials_preceded_own <- mean(preceeded_by_own_trial_stats$count_preceeded_by_own)
range_trials_preceded_own <- range(preceeded_by_own_trial_stats$count_preceeded_by_own)
sd_trials_preceded_own <- sd(preceeded_by_own_trial_stats$count_preceeded_by_own)

print(mean_trials_preceded_own) # 236
print(range_trials_preceded_own) # 144 284
print(sd_trials_preceded_own) # 23

mean_trials_preceded_partner <- mean(preceeded_by_partner_trial_stats$count_preceeded_by_partner)
range_trials_preceded_partner <- range(preceeded_by_partner_trial_stats$count_preceeded_by_partner)
sd_trials_preceded_partner <- sd(preceeded_by_partner_trial_stats$count_preceeded_by_partner)

print(mean_trials_preceded_partner) # 236
print(range_trials_preceded_partner) # 194 261
print(sd_trials_preceded_partner) # 15

# distribution of consecutive active trials

# this part colleague helped with the code
# to show that the design is fully randomized, we can show a distribution of
# consecutive "active" trials per or across participants. this would make clear
# that not one person answered like 5 times in a row or something thus leading
# to some kind of confound in the data.

# this part TS helped me, so I directly import the same data that was revised
# and checked by her

library(readr)
Data <- read_csv("data_checked_by_colleage.csv")

d_check <- Data

# code in python:
# data["CumSum"] = data["comparison_In"].groupby(data["comparison_In"].eq(False).cumsum()).cumsum().tolist()

# create a lag of the IDs
d_check$lagged_id <- lag(d_check$individual_id)

# compare the IDs with the lagged IDs
d_check$comparison <- d_check$individual_id == d_check$lagged_id

## you can convert TRUEs to 1s and 0s for FALSE

# count the cumulative sum (see python code)
d_check$CumSum <- with(d_check, ave(comparison_In, cumsum(!comparison_In), FUN = cumsum))

# Print the count of data points for each unique value of CumSum
d_check %>%
  group_by(CumSum) %>%
  summarise(count = n()) %>%
  print()

# visualize
count_plot <- d_check %>%
  ggplot(aes(x = CumSum)) +
  geom_bar() +
  labs(#title = "Distribution of Consecutive Active Trials",
       x = "Number of Consecutive Trials",
       y = "Count") + theme_pubr(base_size = 14)

print(count_plot)
```

# Active trials count per pair/dayd

```{r}
# Count the number of left/right responses per pair

# For left responses per pair
left_responses_per_pair <- d %>%
  filter(response_n == "0") %>%
  group_by(pair) %>%
  summarise(count_leftward_per_pair = sum(length(trial)))

# For right responses per pair
right_responses_per_pair <- d %>%
  filter(response_n == "1") %>%
  group_by(pair) %>%
  summarise(count_rightward_per_pair = sum(length(trial)))

# Merge left and right counts for each pair
left_right_responses_per_pair <- left_responses_per_pair %>%
  full_join(right_responses_per_pair, by = "pair")

# Summary statistics for left/right responses per pair
mean_left_per_pair <- mean(left_right_responses_per_pair$count_leftward_per_pair, na.rm = TRUE) # 494
range_left_per_pair <- range(left_right_responses_per_pair$count_leftward_per_pair, na.rm = TRUE) # 413 566
sd_left_per_pair <- sd(left_right_responses_per_pair$count_leftward_per_pair, na.rm = TRUE) # 45.36518

mean_right_per_pair <- mean(left_right_responses_per_pair$count_rightward_per_pair, na.rm = TRUE) # 506
range_right_per_pair <- range(left_right_responses_per_pair$count_rightward_per_pair, na.rm = TRUE) # 434 587
sd_right_per_pair <- sd(left_right_responses_per_pair$count_rightward_per_pair, na.rm = TRUE) # 45.36518

# Print results
print(mean_left_per_pair)
print(range_left_per_pair)
print(sd_left_per_pair)

print(mean_right_per_pair)
print(range_right_per_pair)
print(sd_right_per_pair)
```

# Bias values for two members within the same pair

Potentially left biased participants could be paired with right biased participants. This would look like a preference for alternating responses. **This can be checked by calculating the correlation of left/right bias of participants in the same dyad and visualized by a scatter plot.** On one axis you plot the left/right bias of one participant of a dyad and on the other the left/right bias of the other. The resulting cloud could be round (uncorrelated), or elongated (correlated). In case it is correlated or anticorrelated that data might be understood by means of opposing biases within dyads. That's different from our present message

Essentially, we want to check if, in a dyad, one participant's tendency to respond left/right is associated with the other participant's tendency to respond in the opposite or same direction. This would give insights into whether there’s a tendency toward alternating responses between participants (potentially reflecting opposing biases).

```{r}
# Step 1: Calculate proportion of right responses for each participant
# this is literally just counting how many times a person responded right
proportion_right_responses <- d %>%
  group_by(individual_id) %>%
  summarise(
    pair = first(pair),
    proportion_right_responses = sum(response_n_effect_code == 1) / length(trial)
  )
View(proportion_right_responses)

# Step 2: calculate the proportion of rightward moving stimulus
# this is the number of times the stimulus was rightward
# looks fine to me, so left/right is nearly equal amount
# which is how it's supposed to be due to how we designed the experiment

proportion_stimulus_right <- d %>%
  group_by(individual_id) %>%
  summarise(
    pair = first(pair),
    proportion_stimulus_right = sum(stimulus_n == 1) / length(trial)
  )

View(proportion_stimulus_right)

# Step 3: Left join to align both datasets and calculate "left/right bias"
# bias means a tendency to favor one response over the other i.e., the difference
# between the number of rightward-moving stimulus and the right decisions given
# we do this per each participant

# if the value is positive, that means higher tendency to respond right
# if it's negative that means higher tendency to respond left
# i should not expect a big magnitude though

bias_data <- proportion_right_responses %>%
  left_join(proportion_stimulus_right, by = c("individual_id", "pair")) %>%
  mutate(bias = proportion_right_responses - proportion_stimulus_right)

View(bias_data)

# Step 4: Correlate this "right/left bias" between the two participants of the
# dyads

bias_data_combined <- bias_data %>%
  group_by(pair) %>%
  summarise(
    bias_1 = first(bias),
    bias_2 = last(bias)
  )
View(bias_data_combined)

mean(bias_data$bias)
range(bias_data$bias)
sd(bias_data$bias)

correlation <- cor.test(bias_data_combined$bias_1, bias_data_combined$bias_2)
print(correlation)

# no meaningful relationship between the biases of participants within dyads
# conclusion is the following: 
# While there is a slight tendency for opposing biases (indicated by the
# negative correlation), it is not strong enough to draw any reliable conclusion

# Step 5: Visualize by scatterplot
scatterplot <- ggplot(bias_data_combined, aes(x = bias_1, y = bias_2)) +
  geom_point(size = 2) +
  #geom_smooth(method = "lm", se = FALSE) + 
  labs(
    x = "Bias of Participant 1",
    y = "Bias of Participant 2"
  ) +
  theme_pubr(base_size = 14)

print(scatterplot)
ggsave("Fig_3_scatterplot_bias_dyads.pdf", plot = scatterplot, width = 30, height = 25, units = "cm")
```

# Change of RT over block

- does the reaction time (RT) change over the duration of experiment (blocks)? Maybe people started zooning out during the experiment and took a very long time.

```{r message=FALSE, warning=FALSE}
# calculate mean and sd of RT
mean_rt <- mean(d$rt) # measured in seconds; 0.877 sec
sd_rt <- sd(d$rt) # 0.25 sec

by_block <-group_by(d, block, pair, chamber)
by_block <- summarise(by_block, block_mean = mean(rt))

# plot_x <- ggplot(data = by_block, aes(x=as.factor(block), y=block_mean))+
#   geom_point(aes(colour = chamber))+
#   facet_wrap(~pair)+
#   labs(title = "Mean RT across blocks",
#        x = "block number",
#        y = "mean reaction time (s)") + 
#   ylim(0.4,1.6) + apatheme
# 
# plot_x

# # show a simpler figure
# # show the average reaction time (RT) across the 10 blocks without grouping by
# # dyad and chamber
# 
# by_block <- group_by(d, block)
# by_block <- summarise(by_block, block_mean = mean(rt), block_sd = sd(rt))
# 
# plot_x <- ggplot(data = by_block, aes(x = as.factor(block), y = block_mean)) +
#   geom_point() +
#   geom_errorbar(aes(ymin = block_mean - block_sd, ymax = block_mean + block_sd), width = 0.2) +
#   labs(title = "Mean RT across blocks",
#        x = "Block number",
#        y = "Mean RT (s)") +
#   ylim(0.4, 1.6) +
#   apatheme
# 
# plot_x

#by_block <- group_by(d, block)
#by_block <- summarise(by_block, block_mean = mean(rt), block_sd = sd(rt))

d_RT_block_mean <- d %>%
  group_by(pair, chamber, block) %>%
  summarise(mean_RT = mean(rt))%>%
  group_by(block) %>%
  summarise(mean_RT = mean(mean_RT))


plot_rt <- ggplot(d_RT_block_mean, aes(as.factor(block), mean_RT, group = 1)) +
  geom_point() +
  geom_line() +
  labs(#title = "Mean RT across blocks",
       x = "Block number",
       y = "Mean RT (s)") +
  ylim(0, 1.6) + theme_pubr(base_size = 14)
  #theme_minimal_grid(font_size = 15) # use theme_pubr() instead
  #theme_minimal((axis.text = element_text(size = 15),
  #      axis.title = element_text(size = 15),
  #      panel.grid.major = element_line(color = "gray", linetype = "solid"),
  #      panel.grid.minor = element_blank())

print(plot_rt)
#ggsave("EDA-RT-plot.svg", plot = plot_rt, width = 30, height = 20, units = "cm")
```

# Check accuracy
- do they fall within a reasonable range of around 75%? I don't want people to get too good at the task (>90%) nor too bad at the task (<65%)

```{r message=FALSE, warning=FALSE}

#### mean accuracy/performance calculations ####

# averaged accuracy of everyone per block

# calculating the mean accuracy by counting the number of trials where
# correct_n == 1 (indicating a correct response) and dividing it by the total
# number of trials (n()) for each pair and chamber. This approach calculates the
# proportion of correct responses and gives you the mean correct proportion for
# each pair and chamber.

d_accuracy_summary <- d %>%
  group_by(pair, chamber) %>%
  summarise(mean_correct = sum(correct_n == 1) / n()) 

d_accuracy_summary
mean(d_accuracy_summary$mean_correct) # 0.736555; 73.6%
sd(d_accuracy_summary$mean_correct) # 0.058; 5.8%

# there are two people from two different pair having lower accuracy than 0.6
# i don't think i will remove them because they passed the practice and training

below_threshold <- d_accuracy_summary[d_accuracy_summary$mean_correct < 0.6, ]

# each 66 person's mean accuracy by block

d_accuracy_block_summary <- d %>%
  group_by(pair, chamber, block) %>%
  summarise(mean_correct = sum(correct_n == 1) / n()*100)

# everyone's mean accuracy by block with visualization

d_accuracy_block_mean <- d %>%
  group_by(pair, chamber, block) %>%
  summarise(mean_correct = sum(correct_n == 1) / n() * 100) %>%
  group_by(block) %>%
  summarise(mean_accuracy = mean(mean_correct))

# Plotting 1: averaged performance for all throughout the 10 blocks

plot_accuracy_1 <- ggplot() +
  geom_point(data = d_accuracy_block_mean, aes(x = factor(block), y = mean_accuracy)) +
  geom_line(data = d_accuracy_block_mean, aes(x = factor(block), y = mean_accuracy), group = 'block') +
  ylim(60, 80) +
  labs(#title = "Mean performance across blocks",
       x = "Block number",
       y = "Mean accuracy (%)") + theme_pubr(base_size = 14)
  #theme_minimal_grid(font_size = 15) # use theme_pubr() instead
  #theme_classic() +
  #theme(axis.text = element_text(size = 15),
  #      axis.title = element_text(size = 15))

plot_accuracy_1

#ggsave("EDA-plot-accuracy.svg", plot = plot_accuracy_1, width = 30, height = 20, units = "cm")

# Plotting 2: each 66 individual's mean accuracy shown across 10 blocks

plot_accuracy <- ggplot() +
  geom_point(data = d_accuracy_block_summary, aes(x = factor(block), y = mean_correct, alpha = 1/100, colour = interaction(factor(pair), factor(chamber)), group = interaction(factor(pair), factor(chamber)))) +
  geom_line(data = d_accuracy_block_summary, aes(x = factor(block), y = mean_correct, alpha = 1/100, colour = interaction(factor(pair), factor(chamber)), group = interaction(factor(pair), factor(chamber)))) +
  geom_point(data = d_accuracy_block_mean, aes(x = factor(block), y = mean_accuracy, size = 4)) +
  geom_line(data = d_accuracy_block_mean, aes(x = factor(block), y = mean_accuracy, size = 1), group = 'block') +
  labs(title = "Mean Accuracy Change Over Time",
       x = "Block Number",
       y = "Mean Accuracy") +
  apatheme +
  theme(legend.position = "none")

plot_accuracy

```

# coherence levels calculation

- check the coherence level over time i.e., averaged across trials for each participant.
- wanted to include this to see if the ddm is doing its thing adapting against fatigues and learning

```{r}
# Calculate the mean coherence across blocks for all
range(d$coherence) # 0.06835911 0.70896281
mean(d$coherence) # 0.2126868
sd(d$coherence) # 0.08521328

# Plot 1: Calculate mean coherence for everyone per block
d_coherence_block_mean <- d %>%
  group_by(pair, chamber, block) %>%
  summarise(mean_coherence = mean(coherence)) %>%
  group_by(block) %>%
  summarise(mean_coherence_all = mean(mean_coherence))

plot_coherence_mean <- ggplot(d_coherence_block_mean, aes(x = factor(block), y = mean_coherence_all)) +
  geom_point() +
  geom_line(aes(group = 'block')) +  # Set appropriate y-axis limits for coherence values
  labs(#title = "Mean coherence threshold across blocks",
       x = "Block Number",
       y = "Mean coherence") + 
  ylim(0, 0.25) + theme_pubr(base_size = 14)
  #theme_minimal_grid(font_size = 15)
  #theme(axis.text = element_text(size = 15),
  #      axis.title = element_text(size = 15))

print(plot_coherence_mean)

#ggsave("EDA-plot-coherence.svg", plot = plot_coherence_mean, width = 30, height = 20, units = "cm")

# Calculate range, standard deviation, and mean across blocks for everyone
summary_stats <- d_coherence_block_mean %>%
  summarise(
    coherence_range = range(mean_coherence_all),
    coherence_sd = sd(mean_coherence_all),
    coherence_mean = mean(mean_coherence_all)
  )

# Print the summary statistics
print(summary_stats)

# coherence_range: The range of mean coherence values across blocks, calculated as the difference between the maximum and minimum values.
# 
# coherence_sd: The standard deviation of the mean coherence values across blocks, providing a measure of the variability or spread of these values.
# 
# coherence_mean: The mean of the mean coherence values across blocks, giving an average value for coherence across the different blocks.

# Plot 2: calculate mean coherence for each participant per block

d_coherence_block_mean_each <- d %>%
  group_by(pair, chamber, block) %>%
  summarise(mean_coherence_each = mean(coherence))

plot_coherence <- ggplot(d_coherence_block_mean_each, aes(x = factor(block), y = mean_coherence_each, group = interaction(pair, chamber))) +
  geom_point(size = 3, aes(color = interaction(pair, chamber))) +
  geom_line(size = 1, aes(color = interaction(pair, chamber))) +
  ylim(0.15, 0.25) +  # Set appropriate y-axis limits for coherence values
  labs(#title = "Mean coherence threshold across blocks",
       x = "Block Number",
       y = "Mean Coherence") +
  apatheme +
  theme(legend.position = "none")

print(plot_coherence)

# Plot 3: each person's coherence level across blocks
plot_individual_coherence <- ggplot(d, aes(x = factor(block), y = coherence, group = interaction(pair, chamber), color = interaction(pair, chamber))) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(#title = "Individual Coherence Levels Across Blocks",
       x = "Block Number",
       y = "Coherence") +
  ylim(0, 1) +
  apatheme +
  theme(legend.position = "none")

print(plot_individual_coherence)

# Save the plot
#ggsave("Individual-Coherence-Levels.svg", plot = plot_individual_coherence, width = 30, height = 20, units = "cm")

```

# put the plots together

```{r}
library(cowplot)
# Combine the three plots into one figure
label_size <- 20
combined_plot <- cowplot::plot_grid(plot_rt, plot_accuracy_1, plot_coherence_mean,
                           ncol = 1,
                           labels = c("a", "b", "c"),
                           label_size = label_size)
                           #label_x = 0.03,
                           #labels = c("A", "B", "C"))
# Print the combined plot
print(combined_plot)
ggsave("Fig_4_EDA-combined-plots.pdf", plot = combined_plot, width = 30, height = 25, units = "cm")

```

# Correlations between coherence, accuracy, RT

```{r message=FALSE, warning=FALSE}

#### coherence and accuracy ####
d_coherence_block_summary <- d %>%
  group_by(pair, chamber, block) %>%
  summarise(coherence = mean(coherence))

d_coherence_block_summary$accuracy <- d_accuracy_block_summary$mean_correct

plot_correlation_coh_acc <- ggplot(data=d_coherence_block_summary, aes(x=accuracy, y=coherence))+
  geom_point()+
  geom_smooth(method = lm)+
  apatheme+
  theme(legend.position = "none")

plot_correlation_coh_acc

cor.test(d_coherence_block_summary$coherence, d_coherence_block_summary$accuracy)

# a positive correlation suggests that as the coherence of the dots increases
# (making the task easier), participants tend to have higher accuracy

#### coherence and RT ####

plot_correlation_rt_coh <- ggplot(data=d, aes(x=rt,y=coherence))+
  geom_point()+
  geom_smooth(method = lm)+
  apatheme+
  theme(legend.position = "none")

plot_correlation_rt_coh

cor.test(d$coherence, d$rt)

# a small negative relationship between coherence and RT. 
# as the coherence of the dots increases (indicating an easier task),
# participants tend to exhibit faster reaction times

#### accuracy and RT ####

d_rt_block_summary <- d %>%
  group_by(pair, chamber, block) %>%
  summarise(rt = mean(rt))

d_rt_block_summary$accuracy <- d_accuracy_block_summary$mean_correct

plot_correlation_rt_acc <- ggplot(data=d_rt_block_summary, aes(x=rt,y=accuracy))+
  geom_point()+
  geom_smooth(method = lm)+
  apatheme+
  theme(legend.position = "none")

plot_correlation_rt_acc

cor.test(d_rt_block_summary$rt, d_rt_block_summary$accuracy)

# as the coherence of the dots increases (representing an easier task),
# participants tend to have higher accuracy rates.

```


# Participants' repeat probabilities/behavioral strategy (accuracy excluded)

1. averaged across all 66 individuals
- here, we calculated the averaged repeat probability across each individual (total is 66) - taking all the previous trials performed by the same acting agent or not, and calculate the average probability to repeat, this is because i wanted to see the individual variations in terms of history bias i.e., not caring about the social/dyadic influence (yet).

2. averaged across all dyads (maybe skip this for now)

## repeat probability calculation

- had some slight difficulty for plotting using this

```{r}
# Calculate the repeat probabilities for after self and after partner
repeat_prob <- d %>%
  filter(!is.na(last_done_by_self)) %>%
  group_by(individual_id, last_done_by_self) %>%
  summarise(repeat_prob = sum(repeat_prev_resp == 1) / n()) %>%
  mutate(repeat_prob_percentage = repeat_prob * 100)

head(repeat_prob)
```

## repeat probability calculation II

- i checked to visualize using this function with one csv file. therefore the code is fine, and can be applied to all the participants.

```{r}
# Define a function to compute the repeat probability for trials where the 
# previous response was made by the same participant ("own"), and where it was
# not made by the same participant ("partner"). This function helps in analyzing
# whether participants are more likely to repeat their previous response
# based on whomade the last response.

prob <- function(dataframe, participant_n_1){
  
  # Filter the dataframe to include only trials where the previous response was 
  # made by the participant themselves (participant_n_1 == 1) or partner
  # Remove any rows with missing values in the participant_n_1 column.
  dataframe_own <- dataframe[dataframe$participant_n_1 == 1 & !is.na(dataframe$participant_n_1), ]
  dataframe_partner <- dataframe[dataframe$participant_n_1 == -1 & !is.na(dataframe$participant_n_1), ]
  
  # For the trials where the previous response was made by the same participant 
  # calculate the probability of repeating the same response. This 
  # is done by grouping the data by the 'pair' and 'chamber' columns and then 
  # calculating the proportion of trials where the response was repeated 
  # (repeat_prev_resp == 1).
  repeat_prob_own <- dataframe_own %>%
    group_by(pair, chamber) %>%
    summarise(repeat_prob_own = sum(repeat_prev_resp == 1)/n())

  repeat_prob_partner <- dataframe_partner %>%
    group_by(pair, chamber) %>%
    summarise(repeat_prob_partner = sum(repeat_prev_resp == 1)/n())

  return(list(repeat_prob_own = repeat_prob_own, repeat_prob_partner = repeat_prob_partner))
}

# Call the function and summarize the results
repeat_prob_results <- prob(d, 1)

d1 <- data.frame(
  pair = repeat_prob_results$repeat_prob_own$pair,
  chamber = repeat_prob_results$repeat_prob_own$chamber,
  repeat_prob_own = repeat_prob_results$repeat_prob_own$repeat_prob_own,
  repeat_prob_partner = repeat_prob_results$repeat_prob_partner$repeat_prob_partner
)

d1

# some descriptive stats
average_repeat_prob_own <- mean(d1$repeat_prob_own) # 0.5225361
sd_repeat_prob_own <- sd(d1$repeat_prob_own) # 0.07988518
average_repeat_prob_partner <- mean(d1$repeat_prob_partner) # 0.4713948
sd_repeat_prob_partner <- sd(d1$repeat_prob_partner) # 0.04075678

# convert the pair and chamber variables to factors
d1$pair <- factor(d1$pair)
d1$chamber <- factor(d1$chamber)

# create individual_id column
d1$individual_id <- paste(d1$pair, gsub("chamber_", "", d1$chamber), sep = "")

# Reshape the data to long format
d1_long <- tidyr::pivot_longer(d1, cols = c(repeat_prob_own, repeat_prob_partner),
                               names_to = "trial_type", values_to = "repeat_prob")

# Multiply repeat probabilities by 100
d1_long$repeat_prob <- d1_long$repeat_prob * 100

# Rename trial types
d1_long$trial_type <- ifelse(d1_long$trial_type == "repeat_prob_own", "After Own", "After Partner")

# Visualize the updated repeat probabilities
repeat_prob_plot <- ggplot(d1_long, aes(x = individual_id, y = repeat_prob, fill = trial_type)) +
  geom_bar(stat = "identity", position = position_dodge(preserve = 'single')) +
  labs(x = "Individual ID", y = "Repeat Probability (%)", fill = "Trial Type") +
  theme_pubr(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 9),
    legend.title = element_blank(),
    legend.position = "right"
  )

print(repeat_prob_plot)

ggsave("Supp_Fig_1_repeat_prob.pdf", plot = repeat_prob_plot, width = 30, height = 20, units = "cm")
#ggsave("Supp_Fig_1_repeat_prob.svg", plot = repeat_prob_plot, width = 30, height = 20, units = "cm")

# # Visualize the density of repeat probabilities
# ggplot(d1_long, aes(x = repeat_prob, fill = trial_type)) +
#   geom_density(alpha = 0.5) +
#   labs(x = "Repeat Probability (%)", fill = "Trial Type") +
#   apatheme

# the width/thickness represents the spread or variability of the data i.e., 
# fatter means dispersed or spread out across different values of the variable. 
# the width of the curve gives you an idea of the variability or dispersion of
# the repeat probabilities in this case.
```

# Participants' repeat probabilities/behavioral strategy (accuracy included)

```{r}

d <- d %>%
  mutate(repeat_prev_resp = ifelse(repeat_prev_resp == 1, TRUE, FALSE))

# write a function that computes the repeat probability for a subject:
# 1) after incorrect partner trial
# 2) after incorrect own trial
# 3) after correct partner trial
# 4) after correct own trial

repeat_prob <- function(dataframe, participant_n_1, correct_n_1){

  dataframe <- dataframe[which(dataframe["participant_n_1"] == participant_n_1),]
  dataframe <- dataframe[which(dataframe["correct_n_1"] == correct_n_1),]

    if (participant_n_1 == 1 & correct_n_1 == 1) {
      return(dataframe %>% group_by(pair, chamber) %>%
           summarise(repeat_prob_own_correct = mean(((repeat_prev_resp)))))
    }
   if (participant_n_1 == 1 & correct_n_1 == -1) {
      return(dataframe %>% group_by(pair, chamber) %>%
           summarise(repeat_prob_own_error = mean(((repeat_prev_resp)))))
   }
   if (participant_n_1 == -1 & correct_n_1 == 1) {
      return(dataframe %>% group_by(pair, chamber) %>%
           summarise(repeat_prob_partner_correct = mean(((repeat_prev_resp)))))
   }
   if (participant_n_1 == -1 & correct_n_1 == -1) {
      return(dataframe %>% group_by(pair, chamber) %>%
           summarise(repeat_prob_partner_error = mean(((repeat_prev_resp)))))
   }

}

# call the function and summarize the results, then save all in one dataframe d1

d1<- repeat_prob(d, 1, 1)
d1$repeat_prob_own_error <- repeat_prob(d, 1, -1)$repeat_prob_own_error
d1$repeat_prob_partner_correct <- repeat_prob(d, -1, 1)$repeat_prob_partner_correct
d1$repeat_prob_partner_error <- repeat_prob(d, -1, -1)$repeat_prob_partner_error

d1

all_data <- d1 %>%  pivot_longer(!c(pair,chamber), names_to = c("set",".value"),
               names_pattern = "(own|partner)_(correct|error)") %>% mutate(line_index=paste(pair,chamber))

ggplot(all_data,aes(x=error,y=correct)) +
  geom_point(aes(color=set)) +
  geom_line(aes(group=line_index))+
  geom_hline(yintercept = 0.5, color = "black", size =0.5)+
  geom_vline(xintercept = 0.5, color = "black", size =0.5)+
  annotate("text", x = 0.75, y=0.6, label = "Stay")+
  annotate("text", x = 0.4, y=0.6, label = "Win Stay\nLose Switch")+
  annotate("text", x = 0.35, y=0.4, label = "Switch")+
  annotate("text", x = 0.75, y=0.4, label = "Win Switch\nLose Stay")+
  xlab("P(repeat) after error")+
  ylab("P(repeat) after correct")+
  ggtitle("Behavioral Strategy")+
  apatheme

# add variable to highlight upper right box (stay after correct)
all_data = all_data %>% mutate(
     stay= case_when(
         correct>=.5 & error>=.5 ~ 1,
         TRUE ~ 0
     )
)

all_data = all_data %>% mutate(
     switch = case_when(
         correct<=.5 & error<=.5 ~ 1,
         TRUE ~ 0
     )
)

all_data = all_data %>% mutate(
     stay_switch = case_when(
         set == 'own' & lead(switch) == 1 & stay == 1 ~ 1,
         TRUE ~ 0
     )
)

all_data = all_data %>% mutate(
     stay_switch = case_when(
         set == 'partner' & lag(stay_switch) == 1 ~ 1,
         set == 'own' & stay_switch == 1 ~ 1,
         TRUE ~ 0
     )
)

#plot the same as above but highlight (transparency) the data that is in the right upper box

# one_pair <-all_data[all_data$pair==20220518,];
#
ggplot(all_data,aes(x=error,y=correct)) +
  geom_point(aes(color=set, alpha = stay_switch)) +
  geom_line(aes(group=line_index, alpha=stay_switch))+
  geom_hline(yintercept = 0.5, color = "black", size =0.5)+
  geom_vline(xintercept = 0.5, color = "black", size =0.5)+
  annotate("text", x = 0.75, y=0.6, label = "Stay")+
  annotate("text", x = 0.4, y=0.6, label = "Win Stay\nLose Switch")+
  annotate("text", x = 0.35, y=0.4, label = "Switch")+
  annotate("text", x = 0.75, y=0.4, label = "Win Switch\nLose Stay")+
  xlab("P(repeat) after error")+
  ylab("P(repeat) after correct")+
  ggtitle("Behavioral Strategy")+
  apatheme

#People tend to repeat the action that their partner made even though it was incorrect when their partner made it?
#they switch if their partner made a correct decision?
```
